---
title: "CSF Bulk 2. Differential Expression CSF vs PBMC"
author: "Lindsay N Hayes"
date: "`r Sys.Date()`"
format: html
output-dir: "../code"
execute:
  eval: false
  echo: true
  warning: false
  error: false
  message: false
---

### About the Data
Analysis of FACS data from cells derived from 5-6 mL of cerebrospinal fluid (CSF) or 1-2 million peripheral blood cells from the same individuals. All subjects were virally-suppressed people with HIV. CSF-derived cells were processed fresh. PBMCs were analyzed from frozen aliquots. Cells were FACS gated for Live CD45+ cells then sorted for lymphoid (CD3+, CD20+, or CD56+) and (CD3-, CD20-, and CD56-).

```{r, packages, message=FALSE}
library(DESeq2)
library(ggplot2)
library(cowplot)
library(pheatmap)
library(ggrepel)
library(grid)
library(rstatix)
```

## Create Count Matrix and Deseq object from Feature Counts Output
[reference](https://www.biostars.org/p/277316/#9541262)
```{r, count}
DESeqDataSetFromFeatureCounts <- function (sampleTable, directory = ".", design, ignoreRank = FALSE, ...) 
{
    if (missing(design)) 
        stop("design is missing")
    l <- lapply(as.character(sampleTable[, 2]), function(fn) read.table(file.path(directory, fn), skip=2))
    if (!all(sapply(l, function(a) all(a$V1 == l[[1]]$V1)))) 
        stop("Gene IDs (first column) differ between files.")
    tbl <- sapply(l, function(a) a$V7)
    colnames(tbl) <- sampleTable[, 1]
    rownames(tbl) <- l[[1]]$V1
    rownames(sampleTable) <- sampleTable[, 1]
    dds <- DESeqDataSetFromMatrix(countData = tbl, colData = sampleTable[, 
        -(1:2), drop = FALSE], design = design, ignoreRank, ...)
    return(dds)
}

samples <- read.csv("data/samples.csv", stringsAsFactors = TRUE)
genes <- read.csv("data/genes.csv")

dds <- DESeqDataSetFromFeatureCounts(sampleTable = samples, 
                                     directory = "FC/",
                                     design = ~ 0 + group)

# check
dds
colData(dds)
all.equal(rownames(dds), genes$gene_id)
rowData(dds) <- genes

save(dds, file = "data/full_dds.rda")
```
## Create DDS Results Object
```{r, dds, message=FALSE}

keep <- rowSums(counts(dds) >= 10) >= 6
dds <- dds[keep,]

dds$group <- factor(dds$group, levels = c("PBMC_Myel", "PBMC_Lym", "CSF_Myel", "CSF_Lym"))

dds <- DESeq(dds)
res <- results(dds)
resultsNames(dds)

save(dds, genes, samples, res, file = "data/filt_dds_env.rda")
normalized_counts <- counts(dds, normalized=TRUE)
write.csv(normalized_counts, file = "data/normalized_counts.csv")
```
## PCA 
```{r, pca, fig.height=4, fig.width=6}
vsd <- vst(dds, blind=FALSE)
rownames(vsd) <- rowData(dds)$gene_name
m<-plotPCA(vsd, intgroup=c("tissue", "cell"), returnData = TRUE)
ggplot(m, aes(PC1, PC2, color = group)) + geom_point(size = 3) + scale_color_manual(values = c("purple", "royalblue", "magenta", "tomato")) + theme_cowplot() + xlab("PC1: 72% variance") + ylab("PC2: 14% variance")
```

```{r}
sessionInfo()
```

